Gompertzian growth as infinitely correlated non-local growth

### 2023-08-07 arxiv2/new_draft.tex
update capitals

Another framework used to show how the Gompertz can result from an external or ubiquitous stressor is chemical reaction theory. In reaction theory, a set of reactants combine to produce products, possibly in the presence of a catalyst. Morkov (2019) \cite{markov2019reaction} showed that the Gompertz can be interpreted as growth under a catalyst whose effect diminishes at a rate independent of the growth rate of the reactant. In other words, the catalyst can be considered as an external and ubiquitous stressor or catalyst to the reactants, perfectly in line with the ubiquitous stochastic stressor model from Sacher and De Lauro et al. On the other hand, if the logistic growth equation is recast into the reaction network equivalent, the catalyst is diminished by the growth of the reactant, which suggests that the source of growth is internal or at least dependent on the internal state. So even in reaction theory the gompertz model could be seen as a result of a ubiquitous field while the logistic model could be seen as a result of internal interactions.

@@

First consider a system that grows according to logistic growth. If we let the variable $X(t)$ denote the abundance of a quantity in a normalized system where its maximum size is 1, the equation that determines its size as a function of time is
\begin{equation}
  \frac{1}{X}\frac{d}{dt} X(t) = \beta (1 - X(t)).
\end{equation}
This is commonly interpreted as the right hand side representing ``resources available for growth'' which decreases linearly with respect to the relative abundance's growth rate on the left hand side. One way to generalize logistic growth is to introduce a parameter that exponentiates the resource availability in the following way,
\begin{equation}
\label{eq:Rich}
  \frac{1}{X}\frac{d}{dt} X(t) = \beta (1 - X(t)^{\theta}),
\end{equation}
also called $\theta$-logistic growth or Richards' model. Petroni et al. (2020) argued that this exponentiation can be interpreted as non-linear interaction effects where there is a level of co-operation in the microscopic domain. In the limit where the system becomes \textit{maximaly cooperative} or \textit{coherent}, $\theta\rightarrow 0$ and the $\theta$-logistic equation reduces in this limit to the Gompertz growth model,
 \begin{equation}
 \label{eq:Gomp}
  \frac{1}{X}\frac{d}{dt} X(t) = - \beta \ln X(t).
\end{equation}
In this limit, the relative growth goes to infinity as the system approaches $t=0$, which could be seen as a verification of the maximally coherent characteristic of the system. This definition of coherence in the context of the gompertz equation was shown to rigorously correspond to the definition of coherence used in quantum mechanical systems \cite{molski2003coherent} attributed to Glauber (1963) \cite{glauber1963coherent}.

>>>>

Another framework used to show how the Gompertz can result from an external or ubiquitous stressor is chemical reaction theory. In reaction theory, a set of reactants combine to produce products, possibly in the presence of a catalyst. Morkov (2019) \cite{markov2019reaction} showed that the Gompertz can be interpreted as growth under a catalyst whose effect diminishes at a rate independent of the growth rate of the reactant. In other words, the catalyst can be considered as an external and ubiquitous stressor or catalyst to the reactants, perfectly in line with the ubiquitous stochastic stressor model from Sacher and De Lauro et al. On the other hand, if the logistic growth equation is recast into the reaction network equivalent, the catalyst is diminished by the growth of the reactant, which suggests that the source of growth is internal or at least dependent on the internal state. So even in reaction theory the Gompertz model could be seen as a result of a ubiquitous field while the logistic model could be seen as a result of internal interactions.

@@

First consider a system that grows according to logistic growth. If we let the variable $X(t)$ denote the abundance of a quantity in a normalized system where its maximum size is 1, the equation that determines its size as a function of time is
\begin{equation}
  \frac{1}{X}\frac{d}{dt} X(t) = \beta (1 - X(t)).
\end{equation}
This is commonly interpreted as the right hand side representing ``resources available for growth'' which decreases linearly with respect to the relative abundance's growth rate on the left hand side. One way to generalize logistic growth is to introduce a parameter that exponentiates the resource availability in the following way,
\begin{equation}
\label{eq:Rich}
  \frac{1}{X}\frac{d}{dt} X(t) = \beta (1 - X(t)^{\theta}),
\end{equation}
also called $\theta$-logistic growth or Richards' model. Petroni et al. (2020) argued that this exponentiation can be interpreted as non-linear interaction effects where there is a level of co-operation in the microscopic domain. In the limit where the system becomes \textit{maximaly cooperative} or \textit{coherent}, $\theta\rightarrow 0$ and the $\theta$-logistic equation reduces in this limit to the Gompertz growth model,
 \begin{equation}
 \label{eq:Gomp}
  \frac{1}{X}\frac{d}{dt} X(t) = - \beta \ln X(t).
\end{equation}
In this limit, the relative growth goes to infinity as the system approaches $t=0$, which could be seen as a verification of the maximally coherent characteristic of the system. This definition of coherence in the context of the Gompertz equation was shown to rigorously correspond to the definition of coherence used in quantum mechanical systems \cite{molski2003coherent} attributed to Glauber (1963) \cite{glauber1963coherent}.
